### 2012-02-24 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

For every $v \in V$, abbreviation
\begin{itemize}
\item $v.p$ stands for $p_{c(v)}(v)$ if $p_{c(v)} \in \mathcal{P}$.
\item $v.getChildren()$ stands for $\{w | w \in V, (v,w) \in E\}$ which is a set of all children of $v$.
\item $v.getChild(i)$, $i \in \mathbb{N} \cup \{0\}$, stands for $w \in V$ so that $o(v)(v,w) = i + 1$.
\item $v.getChild(class), class \in C$, stands for $w \in V$ so that $(v,w) \in E$ and $c(w) = class$, if $\Bigl | \{u | u \in V, (v,u) \in E, c(u) = class\}\Bigr | = 1$.
\end{itemize}

@@

For instance, assuming $v \in V, c(v) = LiteralNode$, calls $is(v, LiteralNode)$, $is(v, ExprNode)$ return $true$, while call $is(v, ContentNode)$ returns $false$.

@@

The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type, the type is noted. Otherwise, an attempt of the type deduction of the binding expression is made. The deduction of the type from the expression is presented in Algorithm \ref{ALG_func_determineExpressionType}. On input, it takes an expression node and types of local variables that are valid in the expression (also called context of variables or variable context). The presented algorithm is just shortened illustration since the complete version is too long to be presented in this text. Nevertheless, the principle is straightforward. Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_4}), its type can be either determined directly or it depends on its subexpressions. A part of the function's semantic is separated in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.

@@

\subsection{Auxiliary Functions}
Firstly, we define some auxiliary functions used by algorithms in this chapter.

@@

\FORALL{$stepNode \in pathType.stepNodes$}
	\IF{$stepNode.isAxisStep()$}
		\STATE $axisKind$ = $stepNode.getAxisNode().getAxisKind()$
		\IF{$axisKind$ $\neq$ $CHILD$ AND $axisKind$ $\neq$ $DESCENDANT$}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDFOR

@@

\todo[inline]{Algoritmy pouzivaju nejake nedefinovane funkcie, napr stepNode.getAxisNode().}

@@

As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules to classify each occurrence into one of these cases. For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5. \todo{Popisat dovod nizsej vahy (nie je take jasne, ze je to par key-foreign key), pripadne zmenit vahu.}

@@

\todo[inline]{Alg 6.12 FLWOR expressions - pre podmienky na riadkoch 13 a 17 definovat nejake funkcie}

@@

\todo[inline]{Processing function pre join pattern 3.}

@@

\subsection{Analysis of the Join Pattern Occurrences}
For each found join pattern occurrence, the algorithm displayed in Algorithm \ref{ALG_classification_of_join_pattern_occurrences} decides whether it is (O1) or (O2) case, using the rules from the original method and the rule for the join pattern 3 considering it of case (O1) and assigning it with weight of 0.5.

@@

\todo[inline]{Definovat getParent() na Node.}
\todo[inline]{Popisat preco raz bereme target return path types a raz priamo paths.}

\FORALL{$jpOccurrence \in \mathcal{O}$}
    \IF{$jpOccurrence$ is the for join pattern}
        \STATE mark $jpOccurrence$ as case (O1), weight 1
    \ELSIF{$jpOccurrence$ is the join pattern 3}
        \STATE mark $jpOccurrence$ as case (O1), weight 0.5
    \ELSE
        \STATE $secondBindingNode =$ the second binding node from $jpOccurrence$
        \STATE $flworNode = secondBindingNode.getParent().getParent()$
        \STATE $secondVarName = secondBindingNode.varName$
        \STATE $returnPathTypes = getReturnPathTypes(flworNode, secondVarName)$
        \FORALL{$returnPathType \in returnPathTypes$}
        	\IF{$returnPathType.specialFunctionCalls$ contains one of $min$, $max$, $sum$, $avg$}
        		\STATE mark $jpOccurrence$ as case (O1), weight 1 // R2
        		\STATE continue the outmost for cycle with next $jpOccurrence$
        	\ENDIF
        \ENDFOR
        \FORALL{$returnPathType \in returnPathTypes$}
        	\IF{$returnPathType.specialFunctionCalls$ contains $count$}
        		\STATE mark $jpOccurrence$ as case (O1), weight 0.75 // R3
        		\STATE continue the outmost for cycle with next $jpOccurrence$
        	\ENDIF
        \ENDFOR
        \STATE $targetReturnPaths = getTargetReturnPaths(flworNode, secontVarName)$
        \STATE $retPathsNumber = $ number of paths in $targetReturnPaths$
        \IF{$retPathsNumber > 1$}
        	\STATE mark $jpOccurrence$ as case (O2), weight 1 // R4
        \ELSE
            \STATE mark $jpOccurrence$ as case (O2), weight 0.5 // R5
        \ENDIF
    \ENDIF
    \STATE update $jpOccurrence$ in $\mathcal{O}$
\ENDFOR

@@

>>>>

For every $v \in V$, abbreviation
\begin{itemize}
\item $v.p$ stands for $p_{c(v)}(v)$ if $p_{c(v)} \in \mathcal{P}$.
\item $v.getChildren()$ stands for $\{w | w \in V, (v,w) \in E\}$ which is a set of all children of $v$.
\item $v.getChild(i)$, $i \in \mathbb{N} \cup \{0\}$, stands for $w \in V$ so that $o(v)(v,w) = i + 1$.
\item $v.getChild(class), class \in C$, stands for $w \in V$ so that $(v,w) \in E$ and $c(w) = class$, if $\Bigl | \{u | u \in V, (v,u) \in E, c(u) = class\}\Bigr | = 1$.
\item If exists $u \in V$ so that $(u,v) \in E$, $v.getParent()$ stands for that $u$.
\end{itemize}

@@

For instance, assuming $v \in V, c(v) = LiteralNode$, calls $is(v, LiteralNode)$, $is(v, ExprNode)$ returns $true$, while call $is(v, ContentNode)$ returns $false$.

@@

The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type, the type is noted. Otherwise, an attempt of the type deduction of the binding expression is made. The deduction of the type from the expression is presented in Algorithm \ref{ALG_func_determineExpressionType}. On input, it takes an expression node and types of local variables that are valid in the expression (also called context of variables or variable context). The presented algorithm is just shortened illustration since the complete version is too long to be presented in this text. Nevertheless, the principle is straightforward. Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_1}), its type can be either determined directly or it depends on its subexpressions. A part of the function's semantic is separated in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.

@@

\subsection{Auxiliary Functions} \label{step_4_auxiliary_functions}
Firstly, we define some auxiliary functions used by algorithms in this chapter.

@@

\FORALL{$stepNode \in pathType.stepNodes$}
	\IF{$stepNode.isAxisStep$}
		\STATE $axisKind$ = $stepNode.axisNode.getAxisKind()$
		\IF{$axisKind$ $\neq$ $CHILD$ AND $axisKind$ $\neq$ $DESCENDANT$}
			\RETURN false
		\ENDIF
	\ENDIF
\ENDFOR

@@

@@

As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules (R1 - R5) to classify each occurrence into one of these cases. For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5. \todo{Popisat dovod nizsej vahy (nie je take jasne, ze je to par key-foreign key), pripadne zmenit vahu.}

@@

@@

@@

\subsection{Analysis of the Join Pattern Occurrences}
For each found join pattern occurrence, the algorithm displayed in Algorithm \ref{ALG_classification_of_join_pattern_occurrences} and Algorithms \ref{ALG_function_check_r2}, \ref{ALG_function_check_r3}, \ref{ALG_function_use_r4_r5} decides whether it is (O1) or (O2) case, using the rules from the original method and the rule for the join pattern 3 considering it of case (O1) and assigning it with weight of 0.5.

As it can be seen in Algorithm \ref{ALG_classification_of_join_pattern_occurrences}, rules R2 and R3 are applied using instances of $PathType$ while, rules R4 and R5 use instances of $PathExprNode$. This difference is partially explained in definition of the auxiliary functions in Chapter \ref{step_4_auxiliary_functions}. Rules R4 and R5 count target return paths. If they use instances of $PathType$, they can count one path more times (as is described in the mentioned chapter), and thus, give a wrong result.

@@

\FORALL{$jpOccurrence \in \mathcal{O}$}
    \IF{$jpOccurrence$ is the for join pattern}
        \STATE mark $jpOccurrence$ as case (O1), weight 1
    \ELSIF{$jpOccurrence$ is the join pattern 3}
        \STATE mark $jpOccurrence$ as case (O1), weight 0.5
    \ELSE
        \STATE $secondBindingNode =$ the second binding node from $jpOccurrence$
        \STATE $flworNode = secondBindingNode.getParent().getParent()$
        \STATE $secondVarName = secondBindingNode.varName$
        \STATE $returnPathTypes = getReturnPathTypes(flworNode,$ $secondVarName)$
		\IF{$checkR2(returnPathTypes, jpOccurrence)$}
			\STATE continue
		\ENDIF
		\IF{$checkR3(returnPathTypes, jpOccurrence)$}
			\STATE continue
		\ENDIF
		\STATE $targetReturnPaths = getTargetReturnPaths(flworNode,$ $secontVarName)$
		\STATE $useR4R5(returnPaths, jpOccurrence)$
    \ENDIF
    \STATE update $jpOccurrence$ in $\mathcal{O}$
\ENDFOR

@@

\begin{algorithm}
\caption{Function checkR2}
\label{ALG_function_check_r2}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $returnPathTypes$: A list of target return path types. \\
  $jpOccurrence$: An occurrence of a join pattern to classify.
}

\ENSURE True if the $jpOccurrence$ was classified as case (O1) using the R2 rule. False otherwise.

\FORALL{$returnPathType \in returnPathTypes$}
	\IF{$returnPathType.specialFunctionCalls$ contains one of $min$, $max$, $sum$, $avg$}
    	\STATE mark $jpOccurrence$ as case (O1), weight 1 // R2
  		\RETURN true
  	\ENDIF
\ENDFOR

\RETURN false
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Function checkR3}
\label{ALG_function_check_r3}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $returnPathTypes$: A list of target return path types. \\
  $jpOccurrence$: An occurrence of a join pattern to classify.
}

\ENSURE True if the $jpOccurrence$ was classified as case (O1) using the R3 rule. False otherwise.

\FORALL{$returnPathType \in returnPathTypes$}
    \IF{$returnPathType.specialFunctionCalls$ contains $count$}
        \STATE mark $jpOccurrence$ as case (O1), weight 0.75 // R3
        \RETURN true
    \ENDIF
\ENDFOR

\RETURN false
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Function useR4R5}
\label{ALG_function_use_r4_r5}
\begin{algorithmic}[1]
\REQUIRE{\ \\
  $targetReturnPaths$: A list of target return paths. \\
  $jpOccurrence$: An occurrence of a join pattern to classify.
}

\STATE $retPathsNumber = $ number of paths in $targetReturnPaths$
\IF{$retPathsNumber > 1$}
	\STATE mark $jpOccurrence$ as case (O2), weight 1 // R4
\ELSE
	\STATE mark $jpOccurrence$ as case (O2), weight 0.5 // R5
\ENDIF
\end{algorithmic}
\end{algorithm}
